---
title: "Analysis on Inflation"
output: html_document
---


```{r echo=F, warning=F, message=F}
library(dplyr)
library(tidyr)
library(stringr)
library(data.table)
library(DT)
library(plotly)

fpath <- "./data/"

CPI <- read.csv(paste0(fpath, "CPI.csv"))
GDP <- read.csv(paste0(fpath, "GDP.csv"))
GDPgrowth <- read.csv(paste0(fpath, "GDPgrowth.csv"))
Inflation <- read.csv(paste0(fpath, "Inflation.csv"))

houseCost <- read.csv(paste0(fpath, "HouseCost.csv"))
sharePrice <- read.csv(paste0(fpath, "SharePrice.csv"))
```

```{r echo=F}
CPI <- CPI %>%
  gather("Year", "CPI", X1960:X2020, na.rm=T) %>%
  mutate(Year=as.integer(substr(Year, 2, 5)))

GDP <- GDP%>%
  gather("Year", "GDP", X1960:X2020, na.rm=T) %>%
  mutate(Year=as.integer(substr(Year, 2, 5)))

GDPgrowth <- GDPgrowth %>%
  gather("Year", "GDPgrowth", X1960:X2020, na.rm=T) %>%
  mutate(Year=as.integer(substr(Year, 2, 5)))

Inflation <- Inflation %>%
  gather("Year", "Inflation", X1960:X2020, na.rm=T) %>%
  mutate(Year=as.integer(substr(Year, 2, 5)))

houseCost <- houseCost %>%
  filter(SUBJECT == "PRICERENT") %>%
  select(c("LOCATION", "TIME", "Value")) %>%
  filter(nchar(TIME) == 4) %>%
  mutate(Country.Code=LOCATION,
         Year=as.integer(TIME),
         HouseCost=Value) %>%
  select(c("Country.Code", "Year", "HouseCost"))

sharePrice <- sharePrice %>%
  select(c("LOCATION", "TIME", "Value")) %>%
  filter(nchar(TIME) == 4) %>%
  mutate(Country.Code=LOCATION,
         Year=as.integer(TIME),
         SharePrice=Value) %>%
  select(c("Country.Code", "Year", "SharePrice"))
```

```{r echo=F}
df1 <- merge(x=GDP, y=CPI, by=c("Country.Name","Country.Code", "Year"), all.x=T)
df1 <- merge(x=df1, y=GDPgrowth, by=c("Country.Name","Country.Code", "Year"), all.x=T)
df1 <- merge(x=df1, y=Inflation, by=c("Country.Name","Country.Code", "Year"), all.x=T)
df2 <- merge(x=houseCost, y=sharePrice, by=c("Country.Code", "Year"), all.x=T)

df <- merge(x=df1, y=df2, by=c("Country.Code", "Year"), all.x=T)
df <- df %>%
  drop_na()
```


```{r echo=F}
df2020 <- df %>%
  filter(Year == 2020)
```



# Current State of Global Economy {.tabset}

Examining inflation in 2020 with various factor by country. Inflation is calculated as increase in CPI (consumer price index)

From the graphs, there are some interesting patterns

1. Larger economies tend to have lower inflation, which is expected
2. Visually, there appears to be a positive correlation between GDP growth and inflation
3. Housing cost does not seem to have a correlation with inflation, this is likely
   because either housing cost is not normalized to year-over-year increase or simply
   the fact that most CPI metrics do not take housing cost into account.
4. There is also a positive correlation between inflation and share price which is interesting

Finally, the country that stands out the most is Turkey, with the inflation well-ahead of other countries at >12%.

## GDP 
```{r echo=F, warning=F, message=F}
df2020 %>%
  filter(!is.na(Inflation) && !is.na(GDP)) %>%
  plot_ly(x=~GDP, y=~Inflation,
          type='scatter', mode='markers', color=~Country.Name,
          size=~GDP,
          marker=list(sizemode="diameter", opacity=0.5),
          hover_info='text',
          text=~paste("Country: ", Country.Name, "<br>",
                      "GDP: $", round(1e-9*GDP,2), "B")
          ) %>%
  layout(title="Inflation vs GDP",
         yaxis=list(title="Inflation (%)"),
         xaxis=list(title="GDP (USD)"),
         hovermode="compare")
```


## GDP Growth
```{r echo=F, warning=F, message=F}
df2020 %>%
  filter(!is.na(Inflation) && !is.na(GDPgrowth)) %>%
  plot_ly(x=~GDPgrowth, y=~Inflation,
          type='scatter', mode='markers', color=~Country.Name,
          size=~GDP,
          marker=list(sizemode="diameter", opacity=0.5),
          hover_info='text',
          text=~paste("Country: ", Country.Name, "<br>",
                      "GDP: $", round(1e-9*GDP,2), "B")
          ) %>%
  layout(title="Inflation vs GDP Growth",
         yaxis=list(title="Inflation (%)"),
         xaxis=list(title="GDP Growth (%)"),
         hovermode="compare")
```

## House Cost
```{r echo=F, warning=F, message=F}
df2020 %>%
  filter(!is.na(Inflation) && !is.na(HouseCost)) %>%
  plot_ly(x=~HouseCost, y=~Inflation,
          type='scatter', mode='markers', color=~Country.Name,
          size=~GDP,
          marker=list(sizemode="diameter", opacity=0.5),
          hover_info='text',
          text=~paste("Country: ", Country.Name, "<br>",
                      "GDP: $", round(1e-9*GDP,2), "B")
          ) %>%
  layout(title="Inflation vs Housing Cost",
         yaxis=list(title="Inflation (%)"),
         xaxis=list(title="House Cost"),
         hovermode="compare")
```
## Share Price
```{r echo=F, warning=F, message=F}
df2020 %>%
  filter(!is.na(Inflation) && !is.na(SharePrice)) %>%
  plot_ly(x=~SharePrice, y=~Inflation,
          type='scatter', mode='markers', color=~Country.Name,
          size=~GDP,
          marker=list(sizemode="diameter", opacity=0.5),
          hover_info='text',
          text=~paste("Country: ", Country.Name, "<br>",
                      "GDP: $", round(1e-9*GDP,2), "B")
          ) %>%
  layout(title="Inflation vs Share Price",
         yaxis=list(title="Inflation (%)"),
         xaxis=list(title="Share Price"),
         hovermode="compare")
```


