---
title: "Analysis on Inflation"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---


```{r echo=F, warning=F, message=F}
library(dplyr)
library(tidyr)
library(stringr)
library(data.table)
library(DT)
library(ggplot2)
library(plotly)

fpath <- "./data/"

CPI <- read.csv(paste0(fpath, "CPI.csv"))
GDP <- read.csv(paste0(fpath, "GDP.csv"))
GDPgrowth <- read.csv(paste0(fpath, "GDPgrowth.csv"))
Inflation <- read.csv(paste0(fpath, "Inflation.csv"))

houseCost <- read.csv(paste0(fpath, "HouseCost.csv"))
sharePrice <- read.csv(paste0(fpath, "SharePrice.csv"))
```


```{r echo=F}
CPI <- CPI %>%
  gather("Year", "CPI", X1960:X2020, na.rm=T) %>%
  mutate(Year=as.integer(substr(Year, 2, 5)))

GDP <- GDP%>%
  gather("Year", "GDP", X1960:X2020, na.rm=T) %>%
  mutate(Year=as.integer(substr(Year, 2, 5)))

GDPgrowth <- GDPgrowth %>%
  gather("Year", "GDPgrowth", X1960:X2020, na.rm=T) %>%
  mutate(Year=as.integer(substr(Year, 2, 5)))

Inflation <- Inflation %>%
  gather("Year", "Inflation", X1960:X2020, na.rm=T) %>%
  mutate(Year=as.integer(substr(Year, 2, 5)))

houseCost <- houseCost %>%
  filter(SUBJECT == "PRICERENT") %>%
  select(c("LOCATION", "TIME", "Value")) %>%
  filter(nchar(TIME) == 4) %>%
  mutate(Country.Code=LOCATION,
         Year=as.integer(TIME),
         HouseCost=Value) %>%
  select(c("Country.Code", "Year", "HouseCost"))

sharePrice <- sharePrice %>%
  select(c("LOCATION", "TIME", "Value")) %>%
  filter(nchar(TIME) == 4) %>%
  mutate(Country.Code=LOCATION,
         Year=as.integer(TIME),
         SharePrice=Value) %>%
  select(c("Country.Code", "Year", "SharePrice"))
```

```{r echo=F}
df1 <- merge(x=GDP, y=CPI, by=c("Country.Name","Country.Code", "Year"), all.x=T)
df1 <- merge(x=df1, y=GDPgrowth, by=c("Country.Name","Country.Code", "Year"), all.x=T)
df1 <- merge(x=df1, y=Inflation, by=c("Country.Name","Country.Code", "Year"), all.x=T)
df2 <- merge(x=houseCost, y=sharePrice, by=c("Country.Code", "Year"), all.x=T)

df <- merge(x=df1, y=df2, by=c("Country.Code", "Year"), all.x=T)
```


The following tables gives a high-level summary of the data available for different
countries. The top 25 countries are used here, each has at least 20 years of data.

```{r echo=F}
df1 <- df1 %>%
  drop_na()
df2 <- df2 %>%
  drop_na()

tab1 <- df1 %>%
  summarise(
    numEntries=n(),
    numCountries=length(unique(Country.Code)),
    minYear=min(Year),
    maxYear=max(Year)
  )
tab2 <- df2 %>%
  summarise(
    numEntries=n(),
    numCountries=length(unique(Country.Code)),
    minYear=min(Year),
    maxYear=max(Year)
  )

tab1['Source'] <- 'World Bank'
tab2['Source'] <- 'OCED'

tabSummary <- rbind(tab1, tab2)
tabSummary <- tabSummary[,c(5,1,2,3,4)]
knitr::kable(tabSummary, caption="Data Summary")
```

# Data by Country
```{r echo=F}
df <- df %>%
  drop_na()
```

```{r echo=F, message=F}
tab <- 
  data.table(df)[,.(
      minYear=min(Year),
      maxYear=max(Year),
      nYears=length(unique(Year))
    ), by=c("Country.Name", "Country.Code")] %>%
  arrange(across(nYears, desc)) %>%
  head(25)

# knitr::kable(tab, caption="Overall Summary")
datatable(tab, options=list(
  columnDefs = list(list(className = 'dt-center', targets = 5)),
  pageLength=6
))
```


```{r warning=F, echo=F}
dfNew <- df %>%
  filter(Country.Name %in% unique(tab$Country.Name))
```


```{r echo=F}
plt_inflation <-
  ggplot(dfNew, aes(x=Year, y=Inflation, color=Country.Code)) +
  geom_line() + 
  labs(title="Inflation (%) of the 15 Countries",
       xlab="Year",
       ylab="Inflation (%)")

plt_cpi <- 
  ggplot(dfNew, aes(x=Year, y=CPI, color=Country.Code)) +
  geom_line() + 
  labs(title="Consumer Price Index (CPI) of the 15 Countries",
       xlab="Year",
       ylab="CPI")

plt_rentPx <-
  ggplot(dfNew, aes(x=Year, y=HouseCost, color=Country.Code)) +
  geom_line() + 
  labs(title="Normalized Rent Price of the 15 Countries",
       xlab="Year",
       ylab="Rent Price")

plt_sharePx <-
  ggplot(dfNew, aes(x=Year, y=SharePrice, color=Country.Code)) +
  geom_line() + 
  labs(title="Share Price of the 15 Countries",
       xlab="Year",
       ylab="Share Price")

plt_gdp <-
  ggplot(dfNew, aes(x=Year, y=GDP, color=Country.Code)) +
  geom_line() + 
  labs(title="GDP (USD) of the 15 Countries",
       xlab="Year",
       ylab="GDP (USD)")

plt_gdp_growth <-
  ggplot(dfNew, aes(x=Year, y=GDPgrowth, color=Country.Code)) +
  geom_line() + 
  labs(title="GDP Growth (%) of the 15 Countries",
       xlab="Year",
       ylab="GDP Growth (%)")
```


# Key Economic Indicators {.tabset}

It is clear that out of these top countries, US has the largest GDP.
We can also see that in year 2020, most country has highly negative GDP growth, 
with the exception of Ireland which is interesting.

## Inflation
```{r echo=F}
ggplotly(plt_inflation)
```

## CPI
```{r echo=F}
ggplotly(plt_cpi)
```

## Rent Price
```{r echo=F}
ggplotly(plt_rentPx)
```

## Share Price
```{r echo=F}
ggplotly(plt_sharePx)
```

## GDP
```{r echo=F}
ggplotly(plt_gdp)
```

## GDP Growth
```{r echo=F}
ggplotly(plt_gdp_growth)
```






